DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2019-02-25 rebuttal.tex

\begin{mdframed}[style=comment]
Figure 3: What was resolution used for those maps? Does the projection of the eigenvectors depend on the resolution and band-limit?
\end{mdframed}

Those maps have been produced with a resolution of $N_{side}=16$. The information has been added to the paper. (High resolution is costly as diagonalizing the Laplacian matrix scales as $O(N_{pix}^3)$.)

Since we do not have any convergence result yet, we are not sure about how the projection error evolves. However, if there is convergence (and we believe there is), the eigenvectors associated with the lowest $\ell$ probably converge first.

\subsection{}
\begin{mdframed}[style=comment]
Page 7: "That is much more efficient than filtering with spherical harmonics, even though HEALPix was de- signed as an iso-latitude sampling that has a fast spherical transform. That is especially true for smooth kernels which require a low polynomial degree K. Figure 4 compares the speed of low-pass filtering for Gaussian smoothing using the spherical harmonics and the graph-based method presented here. A naive implementation of our method is ten to twenty times faster for Nside = 2048, with K = 20 and K = 5, respectively, than using the spherical harmonic transform at lmax = 3Nside implemented by the highly optimized libsharp [38] library used by HEALPix." This is not so clear. Is the graph method also performing a Gaussian smoothing with the same characteristics? If so, what are those? How precise is the graph convolution approximating the true harmonic-space Gaussian convolution in this case? This is relevant since some approximations (for instance when analyzing cosmic microwave background data) require high-precision Gaussian smoothing. Finally, is lipsharp optimizing Gaussian smoothing in any way?
\end{mdframed}

The experiment whose result is reported in Figure 4 is intended to report filtering speed, independently of the chosen filter (we used a Gaussian for convenience).
While libsharp is probably not optimizing Gaussian smoothing in particular (w.r.t. filtering with any other filter), it is a highly optimized library for SHT.
We don't care about the approximation quality as, in the context of neural networks, filters are learned.
That is, graph filters are not approximating any predefined true harmonic-space filters anyway.
It is however true that replacing spherical filtering (using the SHT) with graph filtering is an interesting prospect (in terms of speed) that might be possible in the future as suggested by Appendix B (Figure B.13).
We feel that, in its current state, graph filtering is probably not a good enough approximation to be usable when high-precision filtering is required.
Nevertheless, theoretical improvements on a potential equivalence (or convergence) of the Laplacian eigenvectors to the spherical harmonics (see our answer to comment 1.4) could potentially make it competitive.

### 2019-02-25 rebuttal.tex

\subsection{}

\begin{mdframed}[style=comment]
I have read with great interest the paper by Perraudin and collaborators presenting a methodology (and the associated code) for performing deep machine learning on the sphere using novel algorithms.

The paper is well-written, and the results are clearly presented. The litterature as well as the advantages and limitations of new and existing algorithms are discussed thoroughly and fairly. I recommend this paper for publication, provided the following comments are addressed.
\end{mdframed}

Dear reviewer, we thank you for your time and thorough comments.
The manuscript improved thanks to your feedback.
You'll find an answer to all the points you mentioned in this rebuttal.

@@

\begin{mdframed}[style=comment]
First of all, I would like to thank the authors for writing such a well thought out paper on such an interesting method.

The paper contains an extremely well described technique for performing convolutions on the sphere by first transforming the sphere into pixel space using graphs and then using graph-based convolutional filters that are radially symmetric. These convolutions are then used in neural networks, with an example showing classification of different cosmological scenarios directly from convergence maps. There are several benefits to this method over previously described neural networks on spheres (such as using spherical harmonics, which is slow, or using 2D CNNs, which has to learn to deal with distortions from the projection). I found the approximation of the graph approaching spherical harmonics extremely fascinating and will certainly looking more into this myself in the future!

I will happily recommend this paper for publication. Firstly though, I would like access to the data (privately rather than via emailing the group) so that I can check the notebooks thoroughly. The notebooks are extremely well written but I would like to run them first. As well as this I think there are a few extra applications which will improve the manuscript further, and make the method more appealing for users. I would recommend that these ideas are at least seriously considered by the authors and hopefully implemented:
\end{mdframed}

\todo{@Tomek: give him access to data}

Dear reviewer, we would like to thank you back for your time, your good words, and your thorough review.
The manuscript substantially improved thanks to your work.
You'll find an answer to all the points you mentioned in this rebuttal.

### 2019-02-26 rebuttal.tex

\todo{@Tomek: give him access to data}

Dear reviewer, we would like to thank you back for your time, your good words, and your thorough review.
The manuscript substantially improved thanks to your work.
You'll find an answer to all the points you mentioned in this rebuttal.

\subsection{}

\begin{mdframed}[style=comment]
The first, and most important is an example using masks. Since the authors mention how their method should in principle work well with masks, especially in comparison to HEALPix and Clebsh-Gordan transform spherical harmonic methods I think that it would be extremely enlightening to present this work, probably in the form of a performance vs. time plot. With this being done, the informational content of a masked version of the cosmological application could easily be included. This would, I'm sure, be attractive to the cosmological community as well as the ML community who are always looking for good ways to deal with masked data.
\end{mdframed}

We agree with the reviewer that an example with an irregular mask is the most missed.
A performance versus time plot would indeed be a great way to present such a comparison.
As the reviewer hints at, this example would be better presented in the light of a comparison to other formulations of spherical CNNs.
While such a comparison would without any doubt be insightful on many aspects (not only for masked data, but also on the role of isotropic filters or the importance of rotation equivariance), it is better carried out on diverse datasets and tasks, and deserves, in our opinion, a follow-up paper to be properly addressed.
The point of this paper was (1) to show that spherical CNNs are a great model for cosmological applications, and (2) that a graph-based spherical CNN has certain undeniable advantages.
As hinted in the conclusion, we are actively working on a more thorough comparison of graph-based and SHT-based spherical CNNs.

On the practical side, a proper comparison requires to either modify the available implementations of \cite{cohen2018sphericalcnn,esteves2017sphericalcnn,kondor2018clebsch} to use the HEALPix sampling, or to transform the cosmological maps to the sampling schemes they support.
Moreover, implementing the pooling operation for irregular masks is not trivial.
All in all, that is a tremendous amount of engineering work that we preferred to defer.

Keep in mind that, while they are not irregular masks, we already use $1/12$, $1/48$, and $1/192$ parts of the sphere for training and inference.
While an SHT-based method needs to consider the whole sphere to work with this data, our graph-based method only considers the used part (and still takes the curvature of the surface into account).
The computational advantage of doing so has been added to Figure 4.
As stated above, the performance comparison will however have to wait our follow-up work.
% Our main motivation was to build a "large" dataset with a relatively low number of cosmological maps.

% In terms of neural network, the available implementation of \cite{cohen2018sphericalcnn,kondor2018clebsch} for spherical CNNs and Clebsh-Gordan CNNs are not scalable enough to be trained on the cosmological maps dataset and do not use the HEALPix sampling. Hence a comparison with these methods requires a different setting, which far beyond the scope of our contribution.

%\nati{We could compare the price of one convolution. What do you think? We can write the theoretical complexity. This is a sensitive point!}

\mdeff{measure the computational time of filtering on a 1/12, 1/48 or 1/192 of the sphere with the graph. Add the result to Figure 4 and compare with filtering of the full sphere (with SHT and graph).}

### 2019-02-26 rebuttal.tex

\todo{Assigned: @Nati, @Michael, @Tomek}

\todo{experiment in prograss by nati}

\nati{\textbf{Very important point!!!}
\begin{itemize}
  \item We can do this comparison...
  \item Since we used 1/12 of the sphere, all the pixel grids we use are somehow squared. I am now using this as a projection. @Tomek, do you know if this is ok? Or do you know how to do the projection in a simple and efficient manner?
  \item I think that traditional 2D CNNs will be as good as our algorithm. We should be very careful in the way we formulate this in our answer and in the paper.
\end{itemize}
}

\mdeff{Agreed that we should do at least one thing they asked for.
\begin{itemize}
	\item Testing on the 1/12 of the sphere is not what he asked for. On one hand, a global projection would exhibit more distortion. On the other, we never use the whole sphere.
	\item Keep in mind that there is no orientation that is consistent for the 12 base faces (i.e., where is up?).
	\item Agreed that the performance should be very similar. Both are doing local computations. And locally (3 to 5 pixels across), the sphere is isomorphic to an Euclidean plane.
	\item The only difference, in my opinion, would be due to the anisotropic filters the 2D CNN has access to.
	\item Not sure we should include this result in the main paper. Hard to justify in the story, especially as a comparison with alternative spherical CNNs is missing. Let's see the results.
\end{itemize}
}

### 2019-02-26 rebuttal.tex

Dear reviewer, we would like to thank you back for your time, your good words, and your thorough review.
The manuscript substantially improved thanks to your work.
You'll find an answer to all the points you mentioned in this rebuttal.

\subsection{}

\begin{mdframed}[style=comment]
The first, and most important is an example using masks. Since the authors mention how their method should in principle work well with masks, especially in comparison to HEALPix and Clebsh-Gordan transform spherical harmonic methods I think that it would be extremely enlightening to present this work, probably in the form of a performance vs. time plot. With this being done, the informational content of a masked version of the cosmological application could easily be included. This would, I'm sure, be attractive to the cosmological community as well as the ML community who are always looking for good ways to deal with masked data.
\end{mdframed}

We agree with the reviewer that an example with an irregular mask is the most missed.
A performance versus time plot would indeed be a great way to present such a comparison.
As the reviewer hints at, this example would be better presented in the light of a comparison to other formulations of spherical CNNs.
While such a comparison would without any doubt be insightful on many aspects (not only for masked data, but also on the role of isotropic filters or the importance of rotation equivariance), it is better carried out on diverse datasets and tasks, and deserves, in our opinion, a follow-up paper to be properly addressed.
The point of this paper was (i) to show that spherical CNNs are a great model for cosmological applications, and (ii) that a graph-based spherical CNN has certain undeniable advantages.
As hinted in the conclusion, we are actively working on a more thorough comparison of graph-based and SHT-based spherical CNNs.

@@

\todo{experiment in prograss by nati}

\nati{\textbf{Very important point!!!}
\begin{itemize}
  \item We can do this comparison...
  \item Since we used 1/12 of the sphere, all the pixel grids we use are somehow squared. I am now using this as a projection. @Tomek, do you know if this is ok? Or do you know how to do the projection in a simple and efficient manner?
  \item I think that traditional 2D CNNs will be as good as our algorithm. We should be very careful in the way we formulate this in our answer and in the paper.
\end{itemize}
}

\mdeff{Agreed that we should do at least one thing they asked for.
\begin{itemize}
	\item Testing on the 1/12 of the sphere is not what he asked for. On one hand, a global projection would exhibit more distortion. On the other, we never use the whole sphere.
	\item Keep in mind that there is no orientation that is consistent for the 12 base faces (i.e., where is up?).
	\item Agreed that the performance should be very similar. Both are doing local computations. And locally (3 to 5 pixels across), the sphere is isomorphic to an Euclidean plane.
	\item The only difference, in my opinion, would be due to the anisotropic filters the 2D CNN has access to.
	\item Not sure we should include this result in the main paper. Hard to justify in the story, especially as a comparison with alternative spherical CNNs is missing. Let's see the results.
\end{itemize}
}

\begin{mdframed}[style=comment]
Similarly, I think it would useful to check how well the DeepSphere CNN does against the FCN when the input data is rotated randomly during training - I imagine that the results should match or probably exceed the FCN results and I think this would be a neat little test which doesn't involve needing extra data, just more augmentation.
\end{mdframed}

That comparison is not warranted.
% or justified. pointless is maybe too aggressive
There are two ways to deal with symmetries one wants to be invariant to: (i) build them into the architecture of the model, or (ii) augment the dataset such that a (more general) model learns them ``brute-force''.
With respect to rotation invariance, the FCN architecture is of the first kind, while the CNN is of the second kind.
Assuming the task is truly invariant to said symmetry,\footnote{As argued in the introduction, cosmological maps are rotation invariant.} the performance of the first kind of models would be exactly the same as without augmentation, while the second kind would only catch up after seeing enough data to learn the invariance.
That is a waste of resources which is only justified if it's not possible to back the desired invariance in the architecture.
We updated the manuscript to more precisely reflect this argumentation.

% \nati{Technically, testing this is complicated as rotation on HEALPix are complicated and expensive. We probably will have to break a lot of code to do it.
% @Michael: what do you think?}
% \mdeff{Agreed.}

### 2019-02-26 scnn.tex
comment 2.3

For tasks that are rotation invariant, FCNs need less training data as the rotation symmetry is backed in the architecture and need not be learned.
Than can be seen as intrinsic data augmentation, as a CNN would need to see many rotated versions of the same data to learn the invariance.
Moreover, FCNs can accommodate maps with a varying number of pixels.
% agnostic to the size of the input data, i.e., it can classify maps which do not have the same number of pixels.
Such global summarization as the $AV$ layer is commonly used along graph convolutions to classify graphs of varying sizes (as it is invariant to node permutation) \citep{duvenaud2015gcn, li2015gatedgnn}.
As such, CNNs should only be used if rotation invariance is undesired.

@@

The simulations are created using the fast lightcone method \pkg{UFalcon} described in \citep{sgier2018fastgeneration}.
A brief overview about the map making procedure used in \pkg{UFalcon} as well as the simulation parameters are given in Appendix~\ref{sec:convergence_mass_maps}.
We however use a single simulation box, as opposed to two used in that work, as we use source galaxies at a lower redshift of $z=0.8$, instead of $z=1.5$.
\pkg{L-PICOLA} \citep{howlett2015lpicola} is used for fast and approximate N-body simulations.
For each class, we generate $30$ simulations.
Out of the $60$ simulations, $20$ are kept as the test set, and $20\%$ of the remaining training data is used as a validation set, to monitor the training process and select the hyper-parameters.
\figref{map_sample} shows the whole sky simulations and a zoom region for both models.
Initial conditions for these simulations are the same, so the differences in structures can only be attributed to different cosmological parameters used to evolve the particle distribution.

>>>>

There are in general two ways to deal with symmetries one wants to be invariant to: (i) build them into the architecture, or (ii) augment the dataset such that a (more general) model learns them.
With respect to rotation invariance, the FCN architecture is of the first kind, while the CNN is of the second kind.
For tasks that are rotation invariant, FCNs hence need less training data as the rotation symmetry is backed in the architecture and need not be learned.
Than can be seen as intrinsic data augmentation, as a CNN would need to see many rotated versions of the same data to learn the invariance.
Moreover, FCNs can accommodate maps with a varying number of pixels.
% agnostic to the size of the input data, i.e., it can classify maps which do not have the same number of pixels.
Such global summarization as the $AV$ layer is commonly used along graph convolutions to classify graphs of varying sizes (as it is invariant to node permutation) \citep{duvenaud2015gcn, li2015gatedgnn}.
As such, CNNs should only be used if rotation invariance is undesired.

@@

The simulations are created using the fast lightcone method \pkg{UFalcon} described in \citep{sgier2018fastgeneration}.
A brief overview about the map making procedure used in \pkg{UFalcon} as well as the simulation parameters are given in \ref{sec:convergence_mass_maps}.
We however use a single simulation box, as opposed to two used in that work, as we use source galaxies at a lower redshift of $z=0.8$, instead of $z=1.5$.
\pkg{L-PICOLA} \citep{howlett2015lpicola} is used for fast and approximate N-body simulations.
For each class, we generate $30$ simulations.
Out of the $60$ simulations, $20$ are kept as the test set, and $20\%$ of the remaining training data is used as a validation set, to monitor the training process and select the hyper-parameters.
\figref{map_sample} shows the whole sky simulations and a zoom region for both models.
Initial conditions for these simulations are the same, so the differences in structures can only be attributed to different cosmological parameters used to evolve the particle distribution.
